Supplementary Materials- A Bayesian networks approach to infer social changes in burials in northeastern Taiwan during the European colonization period

### 2021-05-12 analysis/paper/paper.Rmd
add the raincloud plot to the main text and edit the SI

<!-- burial data and the definition of tie formation  --> 
Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials (details in Supplementary Online Materials). For less frequent prestige goods, including imported ceramics, gold foils, and fish-shaped ornaments, we linked two burials when they both possess each type of goods (i.e. presence or absence). Node attributes here include osteological data, such as age and sex, and cultural data, such as ritual pottery, and a burial value index. Ritual pottery was identified as locally-made ceramics placed above an individual. These pots mostly have sooty vessel exteriors and charred residues on interiors that are interpreted as vessels used for funeral feasting [@Hsieh2009]. We calculated our burial value by summing values of all types of prestige goods in each burial context. For example, for gold-foil beads, we take the total number of burials (90) and divide by number of burials with gold-foil beads (46), to give a value of 90/46 = 1.96. We repeat this for each type of prestige item. For each burial we sum these values of each type of prestige item to get a burial value [@Jorgensen1991].

>>>>

<!-- burial data and the definition of tie formation  --> 
Trade beads are commonly found across burials with substantial differences in quantities, so we described each burial as having one of four levels, high, upper-middle, lower-middle, and low, according to their distributions across all burials (Figure \@ref(fig:raincloud-beads); details in Supplementary Online Materials). For less frequent prestige goods, including imported ceramics, gold foils, and fish-shaped ornaments, we linked two burials when they both possess each type of goods (i.e. presence or absence). Node attributes here include osteological data, such as age and sex, and cultural data, such as ritual pottery, and a burial value index. Ritual pottery was identified as locally-made ceramics placed above an individual. These pots mostly have sooty vessel exteriors and charred residues on interiors that are interpreted as vessels used for funeral feasting [@Hsieh2009]. We calculated our burial value by summing values of all types of prestige goods in each burial context. For example, for gold-foil beads, we take the total number of burials (90) and divide by number of burials with gold-foil beads (46), to give a value of 90/46 = 1.96. We repeat this for each type of prestige item. For each burial we sum these values of each type of prestige item to get a burial value [@Jorgensen1991].

```{r raincloud-beads, fig.cap = "The distributions of different types of trade beads across all burials analyzed in this paper."}
knitr::include_graphics(here::here("analysis", "figures", "000-raincloud-beads.png"))
```

### 2021-05-12 analysis/paper/supplementary-material.Rmd
add the raincloud plot to the main text and edit the SI

<!-- burial data and the definition of tie formation: partially move to SI  -->
For trade beads with substantial differences in quantities between burials, we described each burial as having one of four levels according to their distributions across all burials (Figure \@ref(fig:figure-raincloud-beads)). Gold-foil beads are in levels of high (\>3), upper-middle (3), lower-middle (2), and low (1); carnelian beads are in levels of high (\>6), upper-middle (4-6), lower-middle (3), and low (1-2); glass beads are in levels of high (\>6), upper-middle (3-6), lower-middle (2), and low (1). If burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them. The trade goods we used to define network ties are relatively durable materials, such as glass, stone, and metal, that are less affected by taphonomic factors in buried environments. Despite the exclusion of eight burials lacking clear archaeological contexts, the burials analyzed in this study consisting of 90% Kiwulan burials can robustly represent network structure. Also, each burial has a clear boundary as our analytical unit that reduces sampling errors.

>>>>

<!-- burial data and the definition of tie formation: partially move to SI  -->
For trade beads with substantial differences in quantities between burials, we described each burial as having one of four levels according to their distributions across all burials (Figure \@ref(fig:figure-raincloud-beads), also in the main text). Gold-foil beads are in levels of high (\>3), upper-middle (3), lower-middle (2), and low (1); carnelian beads are in levels of high (\>6), upper-middle (4-6), lower-middle (3), and low (1-2); glass beads are in levels of high (\>6), upper-middle (3-6), lower-middle (2), and low (1). If burial 1 and burial 2 both have high quantities of agate beads, then there will be a tie connecting them. The trade goods we used to define network ties are relatively durable materials, such as glass, stone, and metal, that are less affected by taphonomic factors in buried environments. In addition, each burial has a clear boundary as our analytical unit that reduces sampling errors. Thus, we are confident that the quantities and distributions of beads in burial contexts can reflect the original differences between burials. Despite the exclusion of eight burials lacking clear archaeological contexts, the burials analyzed in this study consisting of 90% Kiwulan burials can effectively represent structures of burial networks.
